Multi-Robot Probabilistic Mapping and Exploration

### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
new prop for entropy, started another appendix which is empty now

@@

\subsection{Shannon's Entropy}
% Description, figure

\subsection{Expected Information Gain}
% Derivation and link to Appendix

@@

>>>>

The goal of autonomous exploration is to select robotic actions designed to minimize map uncertainty, thereby maximizing map information gain. We formulate autonomous exploration as an optimization problem to determine a policy that accounts for map uncertainty and travel cost.

@@

In this section, we define Shannon's entropy as a measure of map uncertainty, and present a novel approach to predict future entropy from an arbitrary ray. Then we discuss computational modifications for real-time implementation.

\subsection{Shannon's Entropy}

Here we present how the probabilistic properties of an occupancy grid map provide key information about the uncertainty of the space for motion planning. Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}. Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
for an individual cell and the entire map, respectively.
Thus, entropy is maximized when $P(\mathbf{m}_i)=0.5$, which corresponds to the largest uncertainty; similarly, entropy is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, which corresponds to the smallest uncertainty. %; thus Shannon's entropy is a measure of uncertainty.

\subsection{Expected Information Gain}

Suppose a future pose candidate and its associated future measurement scan is $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, where $c\in\mathcal C$ such that $\mathcal C=\braces{1,2,...,n_c}$ accounts for all $n_c$ candidates under consideration for the next robot pose. The current pose of the robot is \emph{not} $X_c$ in general, so any change to the probabilistic map from $X_c$ must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity.

\begin{prop}
\label{prop:ExpectedH}
For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where $z_{c,k}$ refers to the distance from $x_c$ to the $k$-th grid cell along the measurement ray. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap} and the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The second term is derived from \refeqn{allEta} with
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.
The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain,
\begin{align}
\label{eqn:expectedInfoGainRay}
\mathcal I(X_c)&=H(P(m))-\text{E}\left[H(P(m|X_c,Z_c))\right].
\end{align}
\end{prop}
\begin{proof}% TODO: add ref?
See Appendix B.
\end{proof}

The proposed approach discretizes the measurement according to assumptions of occupancy grid mapping, and exploits the probabilistic properties uncovered by the proposed inverse sensor model. These are used to directly calculate the expected value of map entropy for a single measurement ray.

@@

%Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

### 2018-06-24 Dissertation/Thesis/tex/pogm.tex
new prop for entropy, started another appendix which is empty now

Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$. Put differently, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray. For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array. This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space, which is easily determined from geometry, described in further detail in Section \ref{sec:RayCasting}. Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1. Later in Section \ref{sec:RayScanComb}, these are also rearranged in computational algorithms.

@@

\begin{prop}
For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times 
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied. The normalizer $\eta_{t,l}$ is given by
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$.
\end{prop}
\begin{proof}% TODO: add ref
See Appendix.
\end{proof}

@@

At the $t$-th time step, consider the measurement scan $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$. One may apply the results of Proposition 1 repeatedly for each ray to obtain a two-dimensional inverse sensor model because
\begin{align}
\label{eqn:RayByRayScanISM}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},...,z_{t,n_z})\nonumber\\&
=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1}|z_{t,2}|...|z_{t,n_z}).
\end{align}
In this approach, each ray is considered individually such that the inverse sensor model updates the probabilities of grid cells from one ray before subsequent rays of $Z_t$ are considered. Therefore, we assume the first measurement ray $z_{t,1}$ is independent of the remaining rays $z_{t,2:n_z}$. For the second ray, namely $z_{t,2}$, this may depend on $z_{t,1}$, but not subsequent rays $z_{t,3:n_z}$. By the time $z_{t,n_z}$ is considered, it may depend on all prior rays $z_{t,1:n_z-1}$. Therefore, the ray-by-ray approach considers rays within a single scan as partially dependent on each other.

>>>>

Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$. Put differently, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray. For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array. This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space, which is easily determined from geometry, described in further detail in Section \ref{sec:RayCasting}. Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition \ref{prop:ISM}. Later in Section \ref{sec:RayScanComb}, these are also rearranged in computational algorithms.

@@

\begin{prop}
\label{prop:ISM}
For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times 
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied. The normalizer $\eta_{t,l}$ is given by
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$.
\end{prop}
\begin{proof}% TODO: add ref?
See Appendix A.
\end{proof}

@@

At the $t$-th time step, consider the measurement scan $Z_t=\{z_{t,1},z_{t,2},...,z_{t,n_z}\}$. One may apply the results of Proposition \ref{prop:ISM} repeatedly for each ray to obtain a two-dimensional inverse sensor model because
\begin{align}
\label{eqn:RayByRayScanISM}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})&%=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1},z_{t,2},...,z_{t,n_z})\nonumber\\&
=P(\mathbf{m}_i|X_{1:t},Z_{1:t-1}|z_{t,1}|z_{t,2}|...|z_{t,n_z}).
\end{align}
In this approach, each ray is considered individually such that the inverse sensor model updates the probabilities of grid cells from one ray before subsequent rays of $Z_t$ are considered. Therefore, we assume the first measurement ray $z_{t,1}$ is independent of the remaining rays $z_{t,2:n_z}$. For the second ray, namely $z_{t,2}$, this may depend on $z_{t,1}$, but not subsequent rays $z_{t,3:n_z}$. By the time $z_{t,n_z}$ is considered, it may depend on all prior rays $z_{t,1:n_z-1}$. Therefore, the ray-by-ray approach considers rays within a single scan as partially dependent on each other.
